Boosted Regression Trees for Predictive Auto-Tuning

### 2011-12-13 paper/draft1.tex

% =============================================================================
\subsection{Learning a Regression Model}

@@

A regression tree is a piece-wise constant function from one vector space to another,
that works by recursively subdividing the input space into constant regions~\citep{breiman, textbook?}.
They are widely used in statistics and data-mining applications because the fitting algorithm is quick and reliable, and the form of the tree can provide insight into the relevant input variables.
We use a standard fitting procedure, which takes a set of
$(x, y)\in \mathbb{R}^k \times \mathbb{R}$ pairs and constructs a tree with a low mean squared error.
To construct each node of a regression tree,
we sort the set $\mathcal{D}$ of ($x,y$) pairs along each of the $k$ features to find the best partitioning $f_{i,\gamma}$ of the input space along feature $i$ at point $\gamma$ (Eqs.~\ref{eq:figamma}, \ref{eq:treeloss}).
\begin{align}
    f_{i,\gamma}(x) &=
    \begin{cases}
        \alpha  &\text{if $x_i < \gamma$} \\
        \beta &\text{if $x_i \geq \gamma$}
    \end{cases}
    \label{eq:figamma}\\
    i^*, \gamma^* &= \operatorname*{argmin}_{i, \gamma} \mathbb{\hat E}\left[ (y - f_{i,\gamma}(x))^2 \right]
    \label{eq:treeloss}
\end{align}
One disadvantage of the regression tree is that it does not make full use of broad patterns in the data -- each partition formed by the fitting procedure is fit independently in the recursive training procedure,
so it impossible for the model to extract more than one bit of information from each training partition (XXX unclear).
This disadvantage is mitigated to a significant extent by the practice of {\em boosting}~\citet{XXX}.

@@

Boosting is an iterative procedure for constructing an {\em ensemble} of regression trees that is coordinated to fit training examples as accurately as possible.
In a recent empirical study of a range of machine learning regressino problems,
boosted decision trees were found to be among the best and easiest models to apply~\cite{caruana:2010}.
On each boosting iteration, a regression tree is fit to the residual error remaining
after all previously-fit models have made their predictions.
There are essentially three parameters that control the boosted regression tree
training procedure:
1) the depth of tree constructed on each boosting iteration,
2) the minimum number of examples to allow at a regression tree leaf, and
3) the number of trees constructed by boosting.
We did not attempt a systematic study of the effect of these variables on performance.
We chose a maximum depth of 4,
a minimum number of examples of 10,
and 100 iterations of boosting.

>>>>

\subsection{Learning a Regression Model}

@@

A regression tree is a piece-wise constant function from one vector space to another,
that works by recursively subdividing the input space into constant regions~\citep{breiman+friedman+olshen+stone:1984, hastie+tibshirani+friedman:2001}.
They are widely used in statistics and data-mining applications because the fitting algorithm is quick and reliable, and the form of the tree can provide insight into the relevant input variables.
We use a standard fitting procedure, which takes a set of
$(x, y)\in \mathbb{R}^k \times \mathbb{R}$ pairs and constructs a tree with a low mean squared error.
To construct each node of a regression tree,
we sort the set $\mathcal{D}$ of ($x,y$) pairs along each of the $k$ features to find the best partitioning $f_{i,\gamma}$ of the input space along feature $i$ at point $\gamma$ (Eqs.~\ref{eq:figamma}, \ref{eq:treeloss}).
\begin{align}
    f_{i,\gamma}(x) &=
    \begin{cases}
        \alpha  &\text{if $x_i < \gamma$} \\
        \beta &\text{if $x_i \geq \gamma$}
    \end{cases}
    \label{eq:figamma}\\
    i^*, \gamma^* &= \operatorname*{argmin}_{i, \gamma} \mathbb{\hat E}\left[ (y - f_{i,\gamma}(x))^2 \right]
    \label{eq:treeloss}
\end{align}
One disadvantage of the regression tree is that it does not make full use of broad patterns in the data -- each partition formed by the fitting procedure is fit independently in the recursive training procedure,
so it impossible for the model to extract more than one bit of information from each training partition.
% XXX unclear
This disadvantage is mitigated to a significant extent by the practice of {\em boosting}.

@@

Boosting is an iterative procedure for constructing an {\em ensemble} of regression trees that is coordinated to fit training examples as accurately as possible.
\citep{schapire:2001,friedman:2002}
In a recent empirical study of a range of machine learning regressino problems,
boosted decision trees were found to be among the best and easiest models to
apply~\citep{caruana+niculescu-mizil:2006}.
On each boosting iteration, a regression tree is fit to the residual error remaining
after all previously-fit models have made their predictions.
There are essentially three parameters that control the boosted regression tree
training procedure:
1) the depth of tree constructed on each boosting iteration,
2) the minimum number of examples to allow at a regression tree leaf, and
3) the number of trees constructed by boosting.
We did not attempt a systematic study of the effect of these variables on performance.
We chose a maximum depth of 4,
a minimum number of examples of 10,
and 100 iterations of boosting.

\subsection{Search Algorithms}

Once an accurate regression model has been fit to the data, it remains to be optimized for novel arguments (Fig.~\ref{fig:autotune_general}, bottom, step 4).
An exhaustive search is the most reliable if it can be afforded, but in our experiments
(as in general) an exhaustive search is prohibitively expensive.
In our experiments we compare two strategies: (1) a hand-chosen grid provided by the authors of the kernel used in our experiments \citep{pinto+cox:2011gcg},
and (2) a generic stochastic hill-climbing search.

% XXX: what is reference?  Why was it chosen?
% DDC: wasn't the reference implementation hand-tuned?
% might also want to say something like:
% Given that parameters were hand-chosen for the reference kernel, no claims are made as to the optimality nor universality of this reference (indeed, different programmers would undoubtedly arrive at different results).  We provide this kernel configuration simply as a reasonable indicator of typical performance made possible by ad hoc experimentation with parameters.

Additionally, \cite{pinto+cox:2011gcg} advocate a particular grid search over
what was estimated to be the most relevant part of the configuration space. This grid iterates over all
combinations of XXX, but leaves the value of XXX at. In our experiments, we call this algorithm the {\em grid} search procedure.
The grid included $72$ points in addition to the reference implementation, for a total of $73$ points.
XXX: isn't 73 prime?? explain this so-called ``grid''.
